---
title: Language Familiarity Modulates Relative Attention to the Eyes and Mouth of a Talker
author: "Timo Klingler"
date: "5/22/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Research Question

Does speech in an unfamiliar language cause adults to attend more to a talker's mouth?

# Hypothesis

We expected that the participants would attend more to the mouth in the unfamiliar than in the familiar language condition.

## Stimuli

- Movie recording of two female models in sound attenuated room and presented on infrared-based eye-tracking system.

- 20 sentences in English, 20 sentences in other language.

- Participants were presented with sequentially presented pairs of video segments, each consisting of the same person audibly uttering a short sentence, followed by an audio-only clip of one of the two sentences. Participants indicated whether the auditory-only clip was extracted from first or second movie by pressing a key on the keyboard.

## Variables

- *Three different datasets for each experiment*
- Participants
- Task
- EyesEnglish
- EyesIcelandic (1,3) / EyesSpanish(1,2)
- MouthEnglish
- MouthIcelandic (1,3) / MouthSpanish (1,2)

# Experiment 1

## Setup

- 60 monolingual English Speakers
  - 30 English / Icelandic
    - 15 begin with unfamiliar Languange, 15 with familiar
  - 30 English / Spanish
    - 15 begin with unfamiliar Languange, 15 with familiar

## Data

```{r, message = FALSE}

require(dplyr)
require(tidyr)

se <- function(x) 
{
  se <- sd(x) / sqrt(length(x))
  return(se)
}

load("/Users/Timo/Downloads/Experiment 1_English_Icelandic.RData")
ex1_ice <- x %>%
  na.omit() %>%
  gather(Type, Value, EnglishEyesWholeProportion:IcelandicMouthWholeProportion)
  
ex1_ice$Type <- gsub("WholeProportion", "", ex1_ice$Type)

ex1_ice <- separate(ex1_ice, Type, into = c("Language", "Region"), sep = "\\B[A-Z]", remove = TRUE)

ex1_ice$Region <- gsub("yes", "Eyes", ex1_ice$Region)
ex1_ice$Region <- gsub("outh", "Mouth", ex1_ice$Region)

ex1_ice_mean <- aggregate(Value ~ Language + Region, ex1_ice, mean)
ex1_ice_se <- aggregate(Value ~ Language + Region, ex1_ice, se)
ex1_ice_agg <- inner_join(ex1_ice_mean, ex1_ice_se, by = c("Language" = "Language", "Region" = "Region"))
names(ex1_ice_agg) <- c("Language", "Region", "Mean", "SE")

load("/Users/Timo/Downloads/Experiment 1_English_Spanish.RData")
ex1_spa <- x %>% 
  na.omit() %>%
  gather(Type, Value, English_Eyes_Whole_Proportion:Spanish_Whole_Mouth_Proportion)

ex1_spa$Type <- gsub("_Whole", "", ex1_spa$Type)
ex1_spa$Type <- gsub("_Proportion", "", ex1_spa$Type)

ex1_spa <- separate(ex1_spa, Type, into = c("Language", "Region"), sep = "\\_", remove = TRUE)

ex1_spa_mean <- aggregate(Value ~ Language + Region, ex1_spa, mean)
ex1_spa_se <- aggregate(Value ~ Language + Region, ex1_spa, se)
ex1_spa_agg <- inner_join(ex1_spa_mean, ex1_spa_se, by = c("Language" = "Language", "Region" = "Region"))
names(ex1_spa_agg) <- c("Language", "Region", "Mean", "SE")

```

## Results

```{r, message = FALSE}

require(ggplot2)
require(ggthemes)

# RGB Blue: 68, 72, 160
# RGB Green: 0, 191, 76

p_ex1_ice <- ggplot(ex1_ice_agg, aes(x = Language, y = Mean, fill = Region)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = Mean - SE, ymax = Mean + SE), 
                position = position_dodge(0.9), width = 0.1, 
                size = 0.5) +
  scale_fill_manual(values = c("#444878", "#00bf4c")) +
  xlab("") +
  ylab("Mean Proportion Fixation Duration (A)") +
  theme_tufte() +
  theme(axis.title.x = element_text(size = 16),
        axis.text.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        axis.text.y = element_text(size = 16),
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 16))
    
p_ex1_ice

p_ex1_spa <- ggplot(ex1_spa_agg, aes(x = Language, y = Mean, fill = Region)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = Mean - SE, ymax = Mean + SE), 
                position = position_dodge(0.9), width = 0.1, 
                size = 0.5) +
  scale_fill_manual(values = c("#444878", "#00bf4c")) +
  xlab("") +
  ylab("Mean Proportion Fixation Duration (B)") +
  theme_tufte() +
  theme(axis.title.x = element_text(size = 16),
        axis.text.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        axis.text.y = element_text(size = 16),
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 16))
    
p_ex1_spa


```

## Statistical Tests

### Overall

```{r, message = FALSE, warning = FALSE}

require(effsize)
require(pwr)

icelandic1_mouth <- filter(ex1_ice, Language == "Icelandic" & Region == "Mouth")
icelandic1_eyes <- filter(ex1_ice, Language == "Icelandic" & Region == "Eyes")
english1_mouth <- filter(ex1_ice, Language == "English" & Region == "Mouth")
english1_eyes <- filter(ex1_ice, Language == "English" & Region == "Eyes")
spanish1_mouth <- filter(ex1_spa, Language == "Spanish" & Region == "Mouth") 
spanish1_eyes <- filter(ex1_spa, Language == "Spanish" & Region == "Eyes") 
english2_mouth <- filter(ex1_spa, Language == "English" & Region == "Mouth")
english2_eyes <- filter(ex1_spa, Language == "English" & Region == "Eyes")

mouth_overall <- c(icelandic1_mouth$Value, spanish1_mouth$Value)
eyes_overall <- c(icelandic1_eyes$Value , spanish1_eyes$Value)

t.test(mouth_overall, eyes_overall, conf.level = 0.999)
cohen.d(mouth_overall, eyes_overall, conf.level = 0.999)
```
  
### Paired Samples: English/Icelandic 
  
```{r}

icelandic1 <- -1* (icelandic1_eyes$Value - icelandic1_mouth$Value)
mean(icelandic1)
sd(icelandic1)

english1 <- -1*(english1_eyes$Value - english1_mouth$Value)
mean(english1)
sd(english1)

t.test(icelandic1, english1, paired = TRUE, conf.level = 0.999)
cohen.d(icelandic1, english1)

```

```{r}

spanish1 <- -1* (spanish1_eyes$Value - spanish1_mouth$Value)
mean(spanish1)
sd(spanish1)

english2 <- -1*(english2_eyes$Value - english2_mouth$Value)
mean(english2)
sd(english2)

t.test(spanish1, english2, paired = TRUE, conf.level = 0.999)
cohen.d(spanish1, english2)

```

## Statistical Power

```{r}
t.power <- function(m1, m2, sd1, sd2, n1, n2, nsim, alpha)
{
  lower <- qt(alpha/2, df = sum(n1, n2) - 2)
  upper <- qt(1 - alpha/2, df = sum(n1, n2) - 2)
  ts <- replicate(nsim,
                 t.test(rnorm(n1, mean = m1, sd = sd1),
                        rnorm(n2, mean = m2, sd = sd2))$statistic)
  
  return(sum(ts < lower | ts > upper) / nsim)
}

t.power(mean(icelandic1), mean(english1), sd(icelandic1), sd(english1),
        length(icelandic1), length(english1), 1000, alpha = 0.001)

t.power(mean(spanish1), mean(english2), sd(spanish1), sd(english2),
        length(spanish1), length(english2), 1000, alpha = 0.001)
```

# Experiment 2

- New cohort of 30 bilingual English / Spanish speakers
- Identical procedure to Experiment 1

```{r}
load("/Users/Timo/Downloads/Experiment 2_English_Spanish_Bilingual.RData")

ex2_spa <- x %>% 
  na.omit() %>%
  gather(Type, Value, English_Eyes_Whole_Proportion:Spanish_Mouth_Whole_Proportion)

ex2_spa$Type <- gsub("_Whole_Proportion", "", ex2_spa$Type)
ex2_spa <- separate(ex2_spa, Type, into = c("Language", "Region"), sep = "\\_", remove = TRUE)

ex2_spa_mean <- aggregate(Value ~ Language + Region, ex2_spa, mean)
ex2_spa_se <- aggregate(Value ~ Language + Region, ex2_spa, se)
ex2_spa_agg <- inner_join(ex2_spa_mean, ex2_spa_se, by = c("Language" = "Language", "Region" = "Region"))
names(ex2_spa_agg) <- c("Language", "Region", "Mean", "SE")
```

## Plot

```{r}
p_ex2_spa <- ggplot(ex2_spa_agg, aes(x = Language, y = Mean, fill = Region)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = Mean - SE, ymax = Mean + SE), 
                position = position_dodge(0.9), width = 0.1, 
                size = 0.5) +
  scale_fill_manual(values = c("#444878", "#00bf4c")) +
  xlab("") +
  ylab("Mean Proportion Fixation Duration (A)") +
  theme_tufte() +
  theme(axis.title.x = element_text(size = 16),
        axis.text.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        axis.text.y = element_text(size = 16),
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 16))
    
p_ex2_spa
```

## Tests

```{r}
spanish2_mouth <- filter(ex2_spa, Language == "Spanish" & Region == "Mouth") 
spanish2_eyes <- filter(ex2_spa, Language == "Spanish" & Region == "Eyes") 
english3_mouth <- filter(ex2_spa, Language == "English" & Region == "Mouth")
english3_eyes <- filter(ex2_spa, Language == "English" & Region == "Eyes")

spanish2 <- -1 * (spanish2_eyes$Value - spanish2_mouth$Value)
mean(spanish2)
sd(spanish2)

english3 <- -1 * (english3_eyes$Value - english3_mouth$Value)
mean(english3)
sd(english3)

t.test(spanish2, english3, conf.level = 0.999, paired = TRUE)
cohen.d(spanish2, english3)
```

## Statistical Power

```{r}

t.power(mean(spanish2), mean(english3), sd(spanish2), sd(english3),
        length(spanish2), length(english3), 1000, alpha = 0.001)

```
## ANOVA

```{r}

require(car)

anov_dat <- rbind(ex1_spa, ex2_spa)
anov_dat$Region <- as.factor(anov_dat$Region)
anov_dat$Condition <- as.factor(anov_dat$Condition)
anov_dat$Language <- as.factor(anov_dat$Language)

mono_eyes <- filter(anov_dat, Condition == "Mono_Task", Region == "Eyes")
mono_mouth <- filter(anov_dat, Condition == "Mono_Task", Region == "Mouth")
mono_diff <- -1 * (mono_eyes$Value - mono_mouth$Value)

bi_eyes <- filter(anov_dat, Condition == "Bi_Task", Region == "Eyes")
bi_mouth <- filter(anov_dat, Condition == "Bi_Task", Region == "Mouth")
bi_diff <- -1 * (bi_eyes$Value - bi_mouth$Value)

anova1_dat <- data.frame(
  Difference = c(spanish1, english2, spanish2, english3),
Language = factor(rep(c("Spanish", "English", "Spanish", "English"), times = c(length(spanish1), length(english2), length(spanish2), length(english3)))),
Expertise = factor(rep(c("Monolingual", "Bilingual"), times = c(length(c(spanish1, english2)), length(c(spanish2, english3))))))

str(anova1_dat)
summary(anova1_dat)
xtabs(Difference ~ Language + Expertise, anova1_dat)

sumary(lm(Difference ~ Expertise*Language, anova1_dat))
```
# Experiment 3

- New cohort of 30 monolingual English speakers
- Identical to Experiment 1 except for the omission of the auditory sequence at the end of the movies. 

```{r}
load("/Users/Timo/Downloads/Experiment 3_English_Icelandic_NoTask.RData")
ex3_ice <- x %>%
  na.omit() %>%
  gather(Type, Value, English_Eyes_Whole_Proportion:Icelandic_Mouth_Whole_Proportion)
  
ex3_ice$Type <- gsub("_Whole_Proportion", "", ex3_ice$Type)

ex3_ice <- separate(ex3_ice, Type, into = c("Language", "Region"), sep = "\\_", remove = TRUE)

ex3_ice_mean <- aggregate(Value ~ Language + Region, ex3_ice, mean)
ex3_ice_se <- aggregate(Value ~ Language + Region, ex3_ice, se)
ex3_ice_agg <- inner_join(ex1_ice_mean, ex3_ice_se, by = c("Language" = "Language", "Region" = "Region"))
names(ex3_ice_agg) <- c("Language", "Region", "Mean", "SE")
```

## Plot

```{r}
p_ex3_ice <- ggplot(ex3_ice_agg, aes(x = Language, y = Mean, fill = Region)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = Mean - SE, ymax = Mean + SE), 
                position = position_dodge(0.9), width = 0.1, 
                size = 0.5) +
  scale_fill_manual(values = c("#444878", "#00bf4c")) +
  xlab("") +
  ylab("Mean Proportion Fixation Duration (B)") +
  theme_tufte() +
  theme(axis.title.x = element_text(size = 16),
        axis.text.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        axis.text.y = element_text(size = 16),
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 16))
    
p_ex3_ice
```

## Tests

```{r}

icelandic2_mouth <- filter(ex3_ice, Language == "Icelandic" & Region == "Mouth")
icelandic2_eyes <- filter(ex3_ice, Language == "Icelandic" & Region == "Eyes")
english4_mouth <- filter(ex3_ice, Language == "English" & Region == "Mouth")
english4_eyes <- filter(ex3_ice, Language == "English" & Region == "Eyes")

icelandic2 <- -1* (icelandic2_eyes$Value - icelandic2_mouth$Value)
mean(icelandic2)
sd(icelandic2)

english4 <- -1*(english4_eyes$Value - english4_mouth$Value)
mean(english4)
sd(english4)

diff_ice2_eng4 <- -1 * (icelandic2 - english4)

t.test(icelandic2, english4, paired = TRUE, conf.level = 0.999)
cohen.d(icelandic2, english4)

```

## Statistical Power

```{r}

t.power(mean(icelandic2), mean(english4), sd(icelandic2), sd(english4),
        length(icelandic2), length(english4), 1000, alpha = 0.001)

```

## ANOVA

```{r}

anov2_dat <- rbind(ex1_ice, ex3_ice)
anov2_dat$Condition <- as.factor(anov2_dat$Condition)
anov2_dat$Language <- as.factor(anov2_dat$Language)

anova(aov(Value ~ Language * Condition, anov2_dat))

```